Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-04 demystify.tex
updates

One main concern for non-CS or non-STEM students is if you have the sufficient technical background, typically obtained through core CS courses.  You will want to show that you have such knowledge through your coursework, projects, or research.
For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your statement.  If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it in your statement.  If you have done research that requires knowledge of Discrete Maths, you can talk about it in your statement.  You can also ask your LoR writers to talk about your technical background.  
In summary, in your application, convince us that you have the technical background to do CS PhD research. 
Note that many online courses teaching about AI does not really help much as these are not core CS knowledge.

@@

\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
No. In fact, student with BS can get MS degree ``along the way'' to PhD.  However, MS can help if it gives research experience or is from a more well-known school than your undergrad institution.

@@

\section{How long does it take to complete the CS PhD program?}\label{sec:time}

@@

\paragraph{Standard Tests \autoref{sec:grades}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS).

\paragraph{Transcripts} You do not need to get your transcripts evaluated by WES or other services (which can be a hassle and inaccurate).  You can just send your official transcripts directly to the school you're applying to.

@@

\paragraph{Research Experiences} You might have more research experiences than international students. For example, you likely have many opportunities to do research as an undergraduates, e.g., through REUs and internships.  You should highlight these experiences in your application.

\paragraph{Standing out} There are very few domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool).  That makes you stand out and can really help your case.

>>>>

One main concern for non-CS or non-STEM students is if you have the sufficient technical background obtained through core CS courses.  You need to show that you have such knowledge through your coursework, projects, or research.
For example, if you have taken a course on Algorithms, even online ones like Coursera, you can talk about it in your SOP.  If you have done a project that requires knowledge of OS or have a professional certification (e.g., A+) through work, you can talk about it.  If you have done research that requires knowledge of Discrete Maths, you can talk about it.  You can also ask your LoR writers to talk about your technical background.  
In summary, in your application, convince us that you have the background to do CS PhD research.

@@

\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
No. In fact, student can get MS degree ``along the way'' to PhD, e.g., after finishing course work in the first 2 years.  However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution.

@@

\section{How long does to complete the CS PhD program?}\label{sec:time}

@@

\paragraph{Standard Tests \autoref{sec:grades}} You do not need to take TOEFL or IELTS because you already did your undergrad (or MS) at a US university.  You might also be more comfortable communicating in English, e.g., contacting professors (\autoref{sec:contact}).

\paragraph{Transcripts} You do not need to get your transcripts evaluated (which can be a hassle and inaccurate).  You can just send your official transcripts directly to the school you're applying to.

@@

\paragraph{Research Experiences} You might have more research experiences than international students. For example, you likely have many opportunities to do research as an undergraduates, e.g., through REUs and internships.  Highlight such experiences in your application.

\paragraph{Standing out} There are very few domestic applications compared to international ones (i.e., domestic students are minority in the CS PhD application pool), and many US universities want to increase the number of domestic students in their programs.
That makes you stand out and can help your case.
